NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-01-22 Manuscript/manuscriptPS.tex
Update manuscript

%\label{}
\section{Introduction}

In NMRlipids I and II project we were looking for a MD model
which would correctly reproduce headgroup and glycerol
backbone structures and cation binding for PC lipid bilayers \cite{botan15,catte16}.
Here we extend the same goal for lipids with negatively charged PS headgroup.
Chemical structure of PS headgroup together with other common biological
lipids is shown in Fig. \ref{lipids}.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/lipids.pdf}
  \caption{\label{lipids}
    Chemical structures and labels for the headgroup carbons.
  }
\end{figure}

@@

\subsection{Solid state NMR experiments}
The experimental protocol is the same used in Ref. \citenum{ferreira16}.

@@

  DPPE  & Slipids \cite{jambeck12b} &0 & 0        & 288 & 9386 & 0  & 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\

@@

Figs. \ref{PShgSIGNSsimpson} and \ref{glycerolNMR} summarize the experimental NMR results for POPS
bilayer sample.

>>>>

%\label{}
\section{Introduction}
Cellular membranes are composed of mixtures of zwitterionic
and negatively charged lipids. Negatively charged lipids are important for
the signalling and lateral organization of cellular membranes \cite{li4}.
Direct interactions of cations and proteins with negatively charged
lipids have been proposed to trigger signaling events \cite{li14,??}.
On the other hand, it has been proposed that negatively charged lipids 
only increase the local concentration of cations in the vicinity of
the membrane and that the binding constant of cations is similar
to zwitterionic and negatively charged lipids \cite{seelig90,??}.

Details of negatively charged lipids and their interactions
with ions and proteins have been experimentally analyzed with
various spectroscopic methods \cite{borle85,macdonald87,roux90,seelig90,pan14,melcrova16,??}
%NMR \cite{seelig90,??}, scattering \cite{pan14,??} and fluorescence spectroscopy
which are often complemented with classical MD simulations \cite{pan14,melcrova16,??}.
The structure of negatively charged lipids and their detailed
interactions with ions or proteins have not, however, been fully resolved
due to the difficulties to interpret the experimental data \cite{??} and
the inaccuracies in simulation models at the
headgroup region \cite{botan15,catte16,ollila16}.

Headgroup and glycerol backbone C-H bond order parameters
calculated from MD simulations have been
recently used to interpret the lipid structures in NMR experiments
and to validate lipid structure and ion binding in simulations of
PC lipid bilayers \cite{botan15,catte16,ollila16,ferreira16}.
In this work we apply the same approach to negatively charged
phosphatidylserine lipids (PS) which compose   
25-35\% of plasma membrane \cite{li14,??}, regulate
surface charge and protein localization \cite{yeung08}, and
induce protein aggregation \cite{zhao04,??}.

%In NMRlipids I and II project we were looking for a MD model
%which would correctly reproduce headgroup and glycerol
%backbone structures and cation binding for PC lipid bilayers \cite{botan15,catte16}.
%Here we extend the same goal for lipids with negatively charged PS headgroup.
%Chemical structure of PS headgroup together with other common biological
%lipids is shown in Fig. \ref{lipids}.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/lipids.pdf}
  \caption{\label{lipids}
    Chemical structures and labels for the headgroup carbons.
  }
\end{figure}

@@

\subsection{Solid state NMR experiments}
The experimental protocol is the same used in Ref. \citenum{ferreira16}.
\todo{Some basic details should be given.}

@@

@@

Figs. \ref{PShgSIGNSsimpson} and \ref{glycerolNMR} summarize the experimental NMR results for POPS
bilayer sample. \todo{New figure, probably combining Figs. \ref{PShgSIGNSsimpson} and \ref{glycerolNMR}, should
be done. This discussion to be finished after this.}
